Faithful Variable Screening for\\ High-Dimensional Convex Regression

### 2015-05-01 tex/scam/reviews-v1/response-v1.tex

\item \rc{At the top
of page 23, the remark regarding mutual incoherence conditions is a
bit confusing. Aren’t the conditions imposed in Theorem 5.1 on the
covariates essentially analogous to the mutual incoherence conditions
in this case? }

The conditions in Theorem 5.1 are indeed analogous and we have edited the remark accordingly. The conditions in Theorem 5.1 are much more strict than the mutual incoherence condition though.

\item \rc{In the simulation section, the authors might consider comparing the
set of variables selected after the AC step only with the set of
variables selected via the entire AC/DC algorithm. This would help
illustrate the necessity of the concave fitting step in addition to
the convex additive fit alone.}

We have added an additional simulation comparing the variables selected via the AC stage and the DC stage. Please see the last paragraph in Section 6.1 as well as Figure 6. The DC stage is slightly helpful in recovering the true variables but its effect is insignificant. We believe that the DC stage is important in theory but not so in practice.

@@

\item \rc{On page 11, after Deônition 3.2, it is written ``when the joint
density satisfies the property that
$\frac{\partial(x_{-j},x_j)}{\partial x_j} = \frac{\partial^2 p(x_{-j},
  x_j)}{\partial x_j^2} = 0$ at boundary points''.
I think the condition ``$p(x_{−j}, x_j ) = 0$ at the boundary points'' is
missing.}

It is not necessary that $p(\mathbf{x}_{-j}, x_j) = 0$ at $x_j = 0,1$; it is necessary that $p(\mathbf{x}_{-j}, x_j) > 0$. We have however, added an additional condition that $p(\mathbf{x})$ has a bounded second derivative. This condition ensures that the marginal density $p(x_j)$ is twice differentiable and $p'(x_j) = p''(x_j) = 0$ at $x_j = 0,1$. With this, it is valid to apply the quotient rule for differentiation to show that boundary flatness holds. Please see the second item in Example 3.2 for the relevant modifications.

\item \rc{On page 12, I did not understand why the functions 
$\frac{\partial^2 p(x_{-j} |  x_j)}{\partial x_j^2}$
and adn 
$\frac{\partial^2 f(x_{-j} |  x_j)}{\partial x_j^2}$
are bounded (they are not supposed to be continuous)}

We indeed need a bounded condition here. We have modified the definition of boundary flatness so that $p(\mathbf{x_{-j}} \,|\, x_j)$ is bounded and has a bounded first and second derivative for all $x_j$ in some open set around the boundary points and for all $\mathbf{x}_{-j}$. Dominated convergenc theorem holds without problems because we only exchange the derivative and the integral at $x_j = 0,1$. We have verified in detail that the domination condition holds in the newly added Section 8.4.1.

@@

We have removed mentions of $p$ here and corrected the constants.

%\item \rc{on page 38, In the sixth paragraph, in the inequality, the sign
%should be $\leq$. In the eighth paragraph, $c''$ must be replaced by $c$. In
%the ninth paragraph, ``Taking an union''.}
%\item \rc{On page 40, on the fifth line, ``empricial''. In the last paragraph,
%``whose size is bounded'' : it is the log of the size.}

\item \rc{I had a problem with the definition of an ε-bracketing. First, in the
definition 8.1, on page 49, `` we define a bracketing'' should be
replaced by `` we define a $\epsilon$-bracketing''. Then it is said that $\rho(f_L ,
f_U ) \leq \epsilon$ and that the corresponding size is $N(\epsilon,C,\rho)$. But then, in
Proposition 8.3 (and almost everywhere else), when $N(2\epsilon,L_2(P),\rho)$ is
used, it is written $\|f_L − f_U \|_{L_2} \leq \epsilon$ (instead of $2\epsilon$. ) (same thing in
Corollary 8.4 for instance : $\epsilon$-bracketing with a size $N(2\epsilon,...)$)}

We have corrected this problem by replacing all instances of $N(2\epsilon,...)$ with $N(\epsilon, ...)$ and updating the bound for $\log N(...)$ accordingly.

%\item \rc{On page 41, problem with $\epsilon$ (same as the previous item).}
\item \rc{On page 40, I think the functions in $G$ are bounded by $2sB$ (not $sB$.).}

This has been noted and corrected across the proofs.

\item \rc{On page 49, in Definition 8.1, $f^U$ and $f^L$ must be replaced by $f_U$ and
$f_L$. On the last line of the page, ``additive convex functions with $s$
components'' : add ``components bounded by $B$''.}

We have corrected the instances in which we erroneously omitted ``... components bounded by $B$''.

\item \rc{On page 49, ``$\int p(x)^2dx \leq (\int p(x)dx)^2$ ``??}

We have corrected this mistake by placing an upper bound $c_u$ on the density $p(\mathbf{x})$ and incorporating that upper bound into this derivation. We now use the correct bound $\sqrt{ \int p(x)^2 dx } \leq c_u$.

%\item \rc{On page 50, a constant 2 is missing in $\epsilon_n$.}
\item \rc{On page 41, in the second paragraph, I think the bound on $\sup_{h_L} |⟨w,h_L⟩_n|$ is
wrong. I would have used the fact that the variables $h_L(X_i) W_i$ are
independent, centered and sub-Gaussian with a scale smaller than
$2\sigma s B$. It gives a bound of order $sB\sigma \sqrt{\frac{\log\frac{2}{\delta}}{n}}$.}

We have updated this derivation.

\item \rc{The term $\epsilon$ supposed to balance the two terms is not correct (for
instance it does not contain $\delta$). There are many small mistakes in the
calculations. In particular, the exponent in $s$ is not the right one.}

The $\epsilon$ we choose is sufficient to get the bounds we want but it does not exactly balance the terms. We left dependence on $\delta$ out of the $\epsilon$ intentionally because we need to verify that $\epsilon$ satisfies certain conditions, see Equation (8.12) for instance. These conditions are much easier to verify if $\epsilon$ does not depend on $\delta$, which we set to be $1/n$ only at the end. We have updated the paper to clearly state that we use a suboptimal $\epsilon$.

\item \rc{In the proof of Theorem 8.4 : too many small mistakes in
calculations, but the final result is OK.}

TODO

\item \rc{On page 45, the term $\epsilon$ supposed to balance the terms is not correct
(same remark about $\delta$). In the last equations, on the first line (bottom
of the page) $n$ is missing in one of the scalar products.}

We also use a suboptimal $\epsilon$ here for convenience. We have updated the paper to clearly state that.

%\item \rc{On page 46: ``taking a union bound and we have that''}
\item \rc{In Lemma 8.3, I do not see why the functions are continuous : the
functions $x_{-k} \rightarrow p(x_{-k} | x_k)$ are not supposed to be continuous .}

The old Lemma 8.3 has been removed. We now just need $x_{-k} \rightarrow p(x_{-k} \,|\, x_k)$ to be bounded, which has been added as a condition. (We now assume for convenience that the joint density $p(\mathbf{x})$ be bounded away from 0 and $\infty$.)

\item \rc{On page 47, in the proof of Lemma 8.4 : notation $\phi_{-j}$
  has not been defined.
In the definition of $A_+$, the sign should be $>$, not $\geq$. Remove ``both'' .}

The old Lemma 8.4 (now Lemma 8.3) has been updated.

%\item \rc{On page 47, ``a sub-exponential random is''.}

### 2015-05-01 tex/scam/additive_faithful.tex
more responses

The boundary flatness condition intuitively states that two conditional densities $p(\mathbf{x}_{-j} \given x_j)$ and $p(\mathbf{x}_{-j} \given x_j')$ are similar when $x_j$ and $x_j'$ are both close to $0$ or both close to $1$. Product densities satisfy boundary flatness trivially. Boundary flatness is a weak condition because it affects only an $\epsilon$-small region around the boundary; $p(\mathbf{x}_{-j} \given x_j)$ can take arbitrary shapes away from the boundary. Boundary flatness also allows arbitrary correlation structure between the variables (provided $p(\mathbf{x}) >0$). In Section~\ref{sec:boundary_flat}, we give a detailed discussion of the boundary flatness condition and show many examples of boundary flat distributions; in particular, we show that any distribution supported on a compact set can be approximated arbitrarily well by boundary flat distributions.

@@

\begin{theorem}
\label{thm:convex_faithful}
Let $p(\mathbf{x})$ be a density supported on $C=[0,1]^p$ and bounded away from $0/\infty$ that satisfies
the boundary flatness property. 
If $f$ is convex and continuously twice differentiable, then $f$ is additively faithful under $p(\mathbf{x})$.
\end{theorem}

@@

\begin{example}
\label{ex:joint_density_flat}
Boundary flatness is satisfied if the joint density becomes flat at the boundary. To be precise, let $p(\mathbf{x})$ be a joint density bounded away from $0/\infty$. Suppose, for all $j$,  $\partial_{x_j} p(x_j, \mathbf{x}_{-j})$ and $\partial^2_{x_j} p(x_j, \mathbf{x}_{-j})$ exist and are bounded for all $\mathbf{x}_{-j}$ and for all $x_j \in [0, \epsilon) \cup (1-\epsilon, 1]$ for some $\epsilon >0$.

Suppose also, for all $j$,
$$
\partial_{x_j} p(x_j,\mathbf{x}_{-j}) =
\partial_{x_j}^2 p(x_j, \mathbf{x}_{-j}) = 0\quad  \trm{at $x_j = 0, 1$,}
$$
then it is straightforward to show that $\partial_{x_j} p(\mathbf{x}_{-j} \given x_j) = \partial_{x_j}^2 p(\mathbf{x}_{-j} \given x_j) = 0$ at $x_j = 0,1$ as well. One need only see that the derivatives of the marginal densities are zero at the boundary and apply the quotient rule.
\end{example}

The next example shows that any bounded density over a hypercube can be approximated arbitrarily well by some boundary flat distributions.

\begin{example}
\label{ex:mixture_approx}
Suppose $p(\mathbf{x})$ is a bounded density over $[\epsilon, 1-\epsilon]^p$ for some $0 < \epsilon < 1/2$. Let $q(\mathbf{x})$ be a boundary flat density over $[0,1]^p$. Define a mixture $p_{\lambda, \epsilon}(\mathbf{x}) = \lambda q(\mathbf{x}) + (1-\lambda) p(\mathbf{x})$ where $0 < \lambda \leq 1$, then $p_{\lambda, \epsilon}(\mathbf{x})$ is boundary flat over $[0,1]^p$. Furthermore, $\epsilon$ and $\lambda$ can be made arbitrarily small and therefore, $p(\mathbf{x})$ can be approximated arbitrarily well (in $L_1$ for example) by a boundary flat distribution. 
\end{example}

The result in Example~\ref{ex:joint_density_flat} further shows that boundary flat densities can arise from monotone marginal transformations of more general random vectors. This gives another way of approximating general densities by boundary flat densities.

\begin{example}
Let $X = (X_1,...,X_p)$ be drawn from a density supported on $[0,1]^p$, bounded away from $0/\infty$, and with a bounded second derivative. Let $\phi : [0,1] \rightarrow [0,1]$ be a monotonically increasing transformation. We will show that under certain conditions on $\phi$, the density of $(\phi(X_1), ..., \phi(X_p))$ is boundary flat.

Suppose $\phi^{-1}$ has a bounded third derivative on some neighborhood around the boundary, say $[0, \epsilon) \cup (1-\epsilon, 1]$ and suppose $\phi^{-1\prime}(z) = \phi^{-1\prime\prime}(z) = \phi^{-1\prime\prime\prime}(z) = 0$ at $z=0,1$.

Then, the density $p(z_1,...,z_p)$ of the transformed variables $z_j = \phi(x_j)$ is 
$$ 
p(z_1,...,z_p) = p_X(\phi^{-1}(z_1), ..., \phi^{-1}(z_p)) \prod_{j=1}^p \phi^{-1\prime}(z_j),
$$ where $p_X(\cdot,...,\cdot)$ is the density of $X$.
One can use the product rule and the chain rule to differentiate $p(z_1,...,z_p)$ and verify that $\partial_{z_j} p(\mathbf{z}) = \partial^2_{z_j} p(\mathbf{z}) = 0$ at $z_j = 0,1$ and thus show that $p(z_1,...,z_p)$ satisfies boundary flatness by Example~\ref{ex:joint_density_flat}.

As a concrete example, we can use the following transformation $\phi_\epsilon$, defined in term of its inverse:
$$\phi_\epsilon^{-1}(z) = \left\{ \begin{array}{cc} z^4 & \trm{if $z \leq \epsilon$} \\ 
                     \frac{1-2\epsilon^4}{1-2\epsilon} z - \frac{ \epsilon + \epsilon^4}{1-2\epsilon} & \trm{if $\epsilon \leq z \leq 1-\epsilon$} \\
                      1-(1-z)^4 & \trm{if $z \geq 1-\epsilon$} \end{array} \right.$$
for any $1/2 > \epsilon > 0$. As $\epsilon$ becomes small, $\phi_\epsilon$ becomes close to the identity and so $(\phi_\epsilon(X_1), ..., \phi_\epsilon(X_p))$ closely approximates the original random variable $(X_1,...,X_p)$. Intuitively, $\phi$ stretches $X$ near the boundary so that changes in the transformed variable $z$ near the boundary corresponds to much smaller change in original variable $x$. 
\end{example}

@@

\begin{theorem}
\label{thm:acdc_faithful}
Suppose $p(\mathbf{x})$ is a density on $C=[0,1]^p$ bounded away from $0/\infty$ that
satisfies the boundary flatness condition. Suppose that $f$ is convex
and continuously twice differentiable. Let $f^*_k$ and $g^*_k$ be as defined in equations
\eqref{eqn:convex_additive} and \eqref{eqn:concave_postprocess}, then the $f^*_k$'s and the $g^*_k$'s are unique. Furthermore, 
$f^*_k = 0$ and $g^*_k = 0$ implies that $\partial_{x_k} f(\mathbf{x})
= 0$, that is, $f$ does not depend on~$x_k$. 
\end{theorem}

@@

\begin{lemma}
\label{cor:faithfulness_extension}
Suppose $p(\mathbf{x})$ is a density on $C=[0,1]^p$ bounded away from $0/\infty$ satisfying the boundary flatness condition.
Let $f(\mathbf{x})$ be a convex and continuously twice differentiable function on $C$. Let $\phi(\mathbf{x}_{-k})$ be a bounded function that does not depend on $x_k$. Then, we have that the unconstrained univariate function
 \begin{align}
h^*_k = \arg\min_{f_k} \mathbb{E} \Bigl[\bigl( f(X) 
           - \phi(X_{-k}) - h_k(X_k) \bigr)^2\Bigr]
\end{align}
is given by $h^*_k(x_k) = \mathbb{E}\bigl[ f(X) - \phi(X_{-k}) \,|\, x_k\bigr]$,
and $ h^*_k = 0$ implies that $\partial_{x_k} f(\mathbf{x}) = 0$.
\end{lemma}

>>>>

The boundary flatness condition intuitively states that two conditional densities $p(\mathbf{x}_{-j} \given x_j)$ and $p(\mathbf{x}_{-j} \given x_j')$ are similar when $x_j$ and $x_j'$ are both close to $0$ or both close to $1$. Product densities satisfy boundary flatness trivially. Boundary flatness is a weak condition because it affects only an $\epsilon$-small region around the boundary; $p(\mathbf{x}_{-j} \given x_j)$ can take arbitrary shapes away from the boundary. Boundary flatness also allows arbitrary correlation structure between the variables (provided $p(\mathbf{x}) >0$). In Section~\ref{sec:boundary_flat}, we give a detailed discussion of the boundary flatness condition and show examples of boundary flat densities; in particular, we show that any density supported on a compact set can be approximated arbitrarily well by boundary flat densities.

@@

\begin{theorem}
\label{thm:convex_faithful}
Let $p(\mathbf{x})$ be a density supported on $C=[0,1]^p$ and bounded away from $0/\infty$ that satisfies the boundary flatness property.

Suppose $f$ is a convex and continuously twice differentiable on an open set containing $C$, then $f$ is additively faithful under $p(\mathbf{x})$.
\end{theorem}

We let the domain of $f$ be slightly larger than $C$ for a technical reason--it is so we can say in the proof that the Hessian of $f$ is positive semidefinite even at the boundary of $C$.

@@

\begin{example}
\label{ex:joint_density_flat}
Boundary flatness is satisfied if the joint density becomes flat at the boundary. To be precise, let $p(\mathbf{x})$ be a joint density bounded away from $0/\infty$ with a bounded second derivative.

Suppose also, for all $j$,
$$
\partial_{x_j} p(x_j,\mathbf{x}_{-j}) =
\partial_{x_j}^2 p(x_j, \mathbf{x}_{-j}) = 0\quad  \trm{at $x_j = 0, 1$,}.
$$
It is then straightforward to show boundary flatness. One can first verify that the derivatives of the marginal density $p(x_j)$ vanishes at $x_j = 0,1$ and then apply the quotient rule on $\frac{p(x_j,\mathbf{x}_{-j}}{p(x_j)}$ to show that $\partial_{x_j} p(\mathbf{x}_{-j} \given x_j) = \partial_{x_j}^2 p(\mathbf{x}_{-j} \given x_j) = 0$ at $x_j = 0,1$ as well. 
\end{example}

The next example shows that any bounded density over a hypercube can be approximated arbitrarily well by boundary flat densities.

\begin{example}
\label{ex:mixture_approx}
Suppose $p_\epsilon(\mathbf{x})$ is a bounded density over $[\epsilon, 1-\epsilon]^p$ for some $0 < \epsilon < 1/2$. Let $q(\mathbf{x})$ be an arbitrary boundary flat density over $[0,1]^p$ (one can take the uniform density for instance). Define a mixture $p_{\lambda, \epsilon}(\mathbf{x}) = \lambda q(\mathbf{x}) + (1-\lambda) p_\epsilon(\mathbf{x})$ where $0 < \lambda \leq 1$, then $p_{\lambda, \epsilon}(\mathbf{x})$ is boundary flat over $[0,1]^p$.

Now, let $p(\mathbf{x})$ be a bounded density over $[0,1]^p$. Let $p_\epsilon(\mathbf{x})$ be the density formed from truncating $p(\mathbf{x})$ in $[\epsilon, 1-\epsilon]^p$. The corresponding mixture $p_{\lambda,\epsilon}(\mathbf{x})$ then approximates $p(\mathbf{x})$ when $\lambda$ and $\epsilon$ are both small.

Since $p_{\lambda,\epsilon}(\mathbf{x})$ remains boundary flat for arbitrarily small $\epsilon$ and $\lambda$, $p(\mathbf{x})$ can be approximated arbitrarily well (in $L_1$ for example) by boundary flat densities. 
\end{example}

% The result in Example~\ref{ex:joint_density_flat} further shows that boundary flat densities can arise from monotone marginal transformations of more general random vectors. This gives another way of approximating general densities by boundary flat densities.

% \begin{example}
% Let $X = (X_1,...,X_p)$ be drawn from a density supported on $[0,1]^p$, bounded away from $0/\infty$, and with a bounded second derivative. Let $\phi : [0,1] \rightarrow [0,1]$ be a monotonically increasing transformation. We will show that under certain conditions on $\phi$, the density of $(\phi(X_1), ..., \phi(X_p))$ is boundary flat.

% Suppose $\phi^{-1}$ has a bounded third derivative on some neighborhood around the boundary, say $[0, \epsilon) \cup (1-\epsilon, 1]$ and suppose $\phi^{-1\prime}(z) = \phi^{-1\prime\prime}(z) = \phi^{-1\prime\prime\prime}(z) = 0$ at $z=0,1$.

% Then, the density $p(z_1,...,z_p)$ of the transformed variables $z_j = \phi(x_j)$ is 
% $$ 
% p(z_1,...,z_p) = p_X(\phi^{-1}(z_1), ..., \phi^{-1}(z_p)) \prod_{j=1}^p \phi^{-1\prime}(z_j),
% $$ where $p_X(\cdot,...,\cdot)$ is the density of $X$.
% One can use the product rule and the chain rule to differentiate $p(z_1,...,z_p)$ and verify that $\partial_{z_j} p(\mathbf{z}) = \partial^2_{z_j} p(\mathbf{z}) = 0$ at $z_j = 0,1$ and thus show that $p(z_1,...,z_p)$ satisfies boundary flatness by Example~\ref{ex:joint_density_flat}.

% As a concrete example, we can use the following transformation $\phi_\epsilon$, defined in term of its inverse:
% $$\phi_\epsilon^{-1}(z) = \left\{ \begin{array}{cc} z^4 & \trm{if $z \leq \epsilon$} \\ 
%                      \frac{1-2\epsilon^4}{1-2\epsilon} z - \frac{ \epsilon + \epsilon^4}{1-2\epsilon} & \trm{if $\epsilon \leq z \leq 1-\epsilon$} \\
%                       1-(1-z)^4 & \trm{if $z \geq 1-\epsilon$} \end{array} \right.$$
% for any $1/2 > \epsilon > 0$. As $\epsilon$ becomes small, $\phi_\epsilon$ becomes close to the identity and so $(\phi_\epsilon(X_1), ..., \phi_\epsilon(X_p))$ closely approximates the original random variables $(X_1,...,X_p)$. Intuitively, $\phi$ stretches $X$ near the boundary so that a change in the transformed variable $z$ near the boundary corresponds to a much smaller change in original variable $x$. 
% \end{example}

@@

\begin{theorem}
\label{thm:acdc_faithful}
Suppose $p(\mathbf{x})$ is a density on $C=[0,1]^p$ bounded away from $0/\infty$ that
satisfies the boundary flatness condition. Suppose that $f$ is convex and continuously twice differentiable on an open set around $C$. Let $f^*_k$ and $g^*_k$ be as defined in equations
\eqref{eqn:convex_additive} and \eqref{eqn:concave_postprocess}, then the $f^*_k$'s and the $g^*_k$'s are unique. Furthermore, 
$f^*_k = 0$ and $g^*_k = 0$ implies that $\partial_{x_k} f(\mathbf{x})
= 0$, that is, $f$ does not depend on~$x_k$. 
\end{theorem}

@@

\begin{lemma}
\label{cor:faithfulness_extension}
Suppose $p(\mathbf{x})$ is a density on $C=[0,1]^p$ bounded away from $0/\infty$ satisfying the boundary flatness condition.
Let $f(\mathbf{x})$ be a convex and continuously twice differentiable function on an open set around $C$. Let $\phi(\mathbf{x}_{-k})$ be a bounded function that does not depend on $x_k$. Then, we have that the unconstrained univariate function
 \begin{align}
h^*_k = \arg\min_{f_k} \mathbb{E} \Bigl[\bigl( f(X) 
           - \phi(X_{-k}) - h_k(X_k) \bigr)^2\Bigr]
\end{align}
is given by $h^*_k(x_k) = \mathbb{E}\bigl[ f(X) - \phi(X_{-k}) \,|\, x_k\bigr]$,
and $ h^*_k = 0$ implies that $\partial_{x_k} f(\mathbf{x}) = 0$.
\end{lemma}
